### 2015-08-11 Paper/paper.tex
Add first ideas and semi-sentences for PFLP paper

\documentclass[
12pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
]{llncs}
% \usepackage[german]{babel}
\title{Declarative Pearl: Lightweight Probabilistic Programming in Curry}
\author{Sandra Dylus}
\institute{Institut f\"ur Informatik, Christian-Albrechts-Universi\"at zu Kiel\\\email{sad@informatik.uni-kiel.de}}
\begin{document}
\maketitle

\begin{abstract}
  Lorem ipsum
\end{abstract}

\section{Introduction}

There are many probabilistic languages around these days; a progress
that started longer than a decade ago, but is currently especially
active. %
These languages have their origin either in functional, logic or
imperative programming, i.e., you can choose a language based on your
favourite paradigm. %
Besides newly developed languages, a few libraries for probabilistic
programming have seen the light of day as well. %
These languages (stand-alone or DSL) all come with a handfull of
useful tools (functions and stuff) to work with probabilistic values
and distributions as well as inference algorithms to solve
probabilistic models that use these values and distributions. %
That said, there is a large subset that all these languages have in
common, dispite their different approaches on semantic, syntax as well
as goals and range of usage, respectively. %
This subset is rather of practical nature and not of big intereset
from a PL researcher perspective: every good library or programming
language needs a handfull of primitives and predefined functions, but
they are not the heart of this piece of software. %

In case of probabilistic programming languages and libraries, we need
to look at the implementation of the actual semantic behaviour -- what
can we express with probabilistic languages and how can we interpret
such an expression? %
What is the intuition behind probability in the context of programming
and what is its meaning? %

Contributions:

\begin{itemize}
\item We present a lightweight library for probabilistic programming
  in Curry. %
  The library's core consists of ... data structures as well as
  ... functions to work on these strucutres, overall the library needs
  less than 30 lines of code. %
\item We can reduce the library code to such a minimum by reusing Curry's
  semantic. %
  The best part is: no further mechanisms are necessary to enable
  probabilistic computations in Curry. %
  That said, we show high connections between current probability
  programming languages and Curry, including lazy evaluation,
  nondeterministic values as well as corresponding built-in search
  abilities and call-time choice. %
\item We present examples that emphasise the benefits of Curry's
  representation of nondeterminsm, which enables early pruning of the
  search space, compared to Haskell's MonadPlus instance for lists,
  which is often used to model nondeterminism. %
\end{itemize}

\section{The Problem}

\section{The Idea}

\begin{itemize}

\item We treat nondeterministic values like \emph{normal} values, that
  is, we can test predicates on such values in order to filter
  unwanted results. %

\begin{verbatim}
filterDist :: (a -> Bool) -> Dist a -> Dist a
filterDist p d | p (value d) = d
\end{verbatim}

  Such a filter is often called \emph{partial identity} and a common
  functional logic design pattern \cite{funcLogPattern}, additionaly,
  it corresponds to notion of \emph{observing} a probabilistic value
  and is a standard primitive in probabilistic programming
  languages. %

\item we model probabilistic values with nondeterminstic values -- a
  probabilistic value is a pair of a value and its probability; a
  probabilistic value %

\begin{verbatim}
biasedCoin = (True, 0.3)
biasedCoin = (False, 0.7)
\end{verbatim}

\begin{verbatim}
coin = (True ? False,0.5)
\end{verbatim}

\item Each nondeterministic choice corresponds to a branch in the
  search tree. %
  When we evaluate the search tree, we start at the root of the search
  tree and go along a path in that tree: we decide for each branch
  which child to take next. %
  Additionally, we keep track of each decision we make in order to
  repeat a decision in case it occurs several times at different
  positions within the search tree. %

\begin{verbatim}
plus :: Dist Int -> Dist Int -> Dist Int
plus dI1 dI2 = (+) <$> dI1 <*> dI2

doubleCoinWithUnexpectedOutcome =
  coin `plus` coin

doubleCoin =
  let c = coin
  in c `plus` c
\end{verbatim}

\end{itemize}

\section{The Details}

\section{Related Work}

\section{Conclusion and Future Work}

\bibliography{/Users/sad/Documents/Papers/used}
\bibliographystyle{plain}
\end{document}